Indirect elicitation as a necessary condition for consistent surrogate losses

### 2020-06-03 neurips-20/cvx-flats.tex
Adding cautionary tales of Gamma being nonempty and writing up Bo's lemma with hopefully small edit distance

If $\gamma$ is a property elicited by a target loss $\ell$, then taking $\propdis$ to be the regret with respect to $\ell$, the two definitions are equivalent. % for the appropriate choice of $\propdis$.
% when using target regret of $r$ (compared to an optimal prediction for $p$) as $\propdis(r,p)$.
\begin{lemma}\label{lem:consistent-loss-implies-prop}
	Given a surrogate loss $L$, link $\psi$, and target loss $\ell$, set
	% Take $\propdis: \R \times \simplex \to \reals_+$ with 
  $\mu(r,p) := R_\ell(r,p)$.
  Then
	$(L, \psi)$ is consistent with respect to $\ell$ if and only if $(L,\psi)$ is $\propdis$-consistent with respect to $\gamma := \prop{\ell}$.
\end{lemma}\jessiet{Slight trickiness here.  Seems like this should still go through if $\Gamma$ can be empty though.  Have $\propdis(r,p) \neq 0$ for all $r$, but could still approach $0$.}
\begin{proof}
	First, observe that $\propdis(r,p) = 0 \iff \exploss{\ell}{r}{p} = \min_{r' \in \R} \exploss{\ell}{r'}{p} \iff r \in \gamma(p)$.\jessie{changed inf to min... right?}
	Now suppose $(L, \psi)$ are consistent with respect to $\ell$, and take any sequence $\{f_m\}$ of measurable hypotheses.
  Rewriting the right-hand side of Definition~\ref{def:consistent-ell},
        \begin{align}
	  &\; \E_D \ell(\psi \circ f_m(X), Y)\to \inf_f \E_D \ell(\psi \circ f(X), Y)   \label{eqn:cons-loss-cond} \\
	  &\iff \E_X R_\ell(\psi \circ f_m(X), D_X) \to 0                               \nonumber  \\
	  &\iff \E_X \propdis(\psi \circ f_m(X), D_X) \to 0~.~                          \label{eqn:cons-prop-cond}
	\end{align}
        Therefore, $\mathbb{E}_D L(f_m(X),Y) \to \inf_f \mathbb{E}_D L(f(X),Y)$ implies (\ref{eqn:cons-loss-cond}) if and only if it implies (\ref{eqn:cons-prop-cond}).
\end{proof}

@@

\begin{theorem}\label{thm:consistent-implies-indir-elic}
	If a surrogate and link pair $(L, \psi)$ is consistent with respect to a property $\gamma$ or loss $\ell$ eliciting $\gamma$, then $(L, \psi)$ indirectly elicits $\gamma$.
\end{theorem}
\begin{proof}
%  \raft{Maybe ref the consistency definition here explicitly, at the top and the end}
	By Lemma~\ref{lem:consistent-loss-implies-prop}, it suffices to show the result for consistency with respect to a property $\gamma$, setting $\gamma := \prop{\ell}$ if $\ell$ is given instead.\jessiet{flagging; but calls on lemma 1 for the technical backbone.}
	We show the contrapositive; suppose $(L, \psi)$ does not indirectly elicit $\gamma$, meaning we have some $p \in \simplex$ so that $u \in \Gamma(p)$ but $\psi(u) \not \in \gamma(p)$, where $\Gamma := \prop{L}$.
	Consider the constant sequence $\{f_m\}$ with $f_m(x) = u$ for all $m,x$, and take $D$ with full support on some $x \in \X$, and let $D_x = p$.
  Since $u \in \Gamma(p)$, we observe $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$; in particular $\E_D L(f_m(X), Y) \to \inf_f L(f(X),Y)$.
%  \raft{Maybe clearer to first say $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$, and then the easy implication of the convergence \jessie{Done}}
	However, we have $\E_X \propdis(\psi \circ f_m(X), D_X) = \propdis(\psi(u_m), p) = \propdis(\psi(u), p) \neq 0$, since $\psi(u) \not \in \gamma(p)$.
%  \raft{Almost -- just clarify that this is the expectation over $D$ too by construction \jessie{Kinda done?}}
	Thus, we observe $(L, \psi)$ is not consistent with respect to $\gamma$ (Definition~\ref{def:consistent-prop}).
\end{proof}

@@

@@

\begin{theorem}\label{thm:cvx-flats}
	If there is a convex surrogate loss $L : \reals^d \times\Y \to \reals$ indirectly eliciting $\gamma$, then for all $p \in \simplex$ and $r \in \R$ such that $r \in \gamma(p)$, there is a flat $F \subseteq \affhull(\simplex)$ of codimension at most $d$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$.
\end{theorem}
\begin{proof}
	Let $\Gamma := \prop{L}$ and suppose $(L, \psi)$ indirectly elicits $\gamma$ for some link $\psi : \reals^d \to \R$, so that for all $u \in \reals^d$, we have $\Gamma_u \subseteq \gamma_{\psi(u)}$.
	Fix $p \in \simplex$ and $r \in \gamma(p)$.
  \jessie{Show that there is some $u \in \Gamma(p)$ such that $\psi(u) = r$.}
	We will show $F \cap \simplex \subseteq \Gamma_u$; the result then follows from $\Gamma_u \subseteq \gamma_r$.
	As $L$ is convex and elicits $\Gamma$, we have $u \in \Gamma(p) \iff \vec 0 \in \partial \exploss{L}{u}{p} \iff \vec 0 \in \oplus_y p_y \partial L(u,y)$, where $\oplus$ denotes the Minkowski sum~\citep[Theorem 4.1.1]{hiriart2012fundamentals}.
  % (Observe that in order for $L$ to be convex, $L(\cdot, y)$ must be convex for all $y \in \Y$.)

>>>>

If $\gamma$ is a property elicited by a target loss $\ell$, then taking $\propdis$ to be the regret with respect to $\ell$, the two definitions are equivalent. % for the appropriate choice of $\propdis$.
% when using target regret of $r$ (compared to an optimal prediction for $p$) as $\propdis(r,p)$.
\begin{lemma}\label{lem:consistent-loss-implies-prop}
	Given a surrogate loss $L$, link $\psi$, and target loss $\ell$, set
	% Take $\propdis: \R \times \simplex \to \reals_+$ with 
  $\mu(r,p) := R_\ell(r,p)$.
  Then
	$(L, \psi)$ is consistent with respect to $\ell$ if and only if $(L,\psi)$ is $\propdis$-consistent with respect to $\gamma := \prop{\ell}$.
\end{lemma}\btw{Don't \emph{need} $\prop{\ell}$ nonempty/nondegenerate here.}
\begin{proof}
	First, observe that $\propdis(r,p) = 0 \iff \exploss{\ell}{r}{p} = \inf_{r' \in \R} \exploss{\ell}{r'}{p} \iff r \in \gamma(p)$.
	Now suppose $(L, \psi)$ are consistent with respect to $\ell$, and take any sequence $\{f_m\}$ of measurable hypotheses.
  Rewriting the right-hand side of Definition~\ref{def:consistent-ell},
        \begin{align}
	  &\; \E_D \ell(\psi \circ f_m(X), Y)\to \inf_f \E_D \ell(\psi \circ f(X), Y)   \label{eqn:cons-loss-cond} \\
	  &\iff \E_X R_\ell(\psi \circ f_m(X), D_X) \to 0                               \nonumber  \\
	  &\iff \E_X \propdis(\psi \circ f_m(X), D_X) \to 0~.~                          \label{eqn:cons-prop-cond}
	\end{align}
        Therefore, $\mathbb{E}_D L(f_m(X),Y) \to \inf_f \mathbb{E}_D L(f(X),Y)$ implies (\ref{eqn:cons-loss-cond}) if and only if it implies (\ref{eqn:cons-prop-cond}).
\end{proof}

@@

\begin{theorem}\label{thm:consistent-implies-indir-elic}
	If a surrogate and link pair $(L, \psi)$ is consistent with respect to a property $\gamma$ or loss $\ell$ eliciting $\gamma$, then $(L, \psi)$ indirectly elicits $\gamma$.
\end{theorem}
\begin{proof}
%  \raft{Maybe ref the consistency definition here explicitly, at the top and the end}
	By Lemma~\ref{lem:consistent-loss-implies-prop}, it suffices to show the result for consistency with respect to a property $\gamma$, setting $\gamma := \prop{\ell}$ if $\ell$ is given instead.
	We show the contrapositive; suppose $(L, \psi)$ does not indirectly elicit $\gamma$, meaning we have some $p \in \simplex$ so that $u \in \Gamma(p)$ but $\psi(u) \not \in \gamma(p)$, where $\Gamma := \prop{L}$.
	Observe that we use the fact $\Gamma(p) \neq \emptyset$.
	Consider the constant sequence $\{f_m\}$ with $f_m(x) = u$ for all $m,x$, and take $D$ with full support on some $x \in \X$, and let $D_x = p$.
  Since $u \in \Gamma(p)$, we observe $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$; in particular $\E_D L(f_m(X), Y) \to \inf_f L(f(X),Y)$.
%  \raft{Maybe clearer to first say $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$, and then the easy implication of the convergence \jessie{Done}}
	However, we have $\E_X \propdis(\psi \circ f_m(X), D_X) = \propdis(\psi(u_m), p) = \propdis(\psi(u), p) \neq 0$, since $\psi(u) \not \in \gamma(p)$.
%  \raft{Almost -- just clarify that this is the expectation over $D$ too by construction \jessie{Kinda done?}}
	Thus, we observe $(L, \psi)$ is not consistent with respect to $\gamma$ (Definition~\ref{def:consistent-prop}).
\end{proof}

@@

\begin{lemma}\label{lem:set-valued-prop-flats}
	Suppose we are given an elicitable property $\gamma : \simplex \toto \R$ and distribution $p \in \simplex$\jessie{relint?} with reports $r,r' \in \R$ with $p \in \gamma_r \cap \gamma_{r'}$.
	Then for any flat $F$ containing $p$, $F \subseteq \gamma_r \implies F \subseteq \gamma_{r'}$.
\end{lemma}
\begin{proof}

\end{proof}

@@

\begin{theorem}\label{thm:cvx-flats}
Let $\gamma, p, r$ be given, with $p \in \gamma_r$ and either $\gamma$ elicitable or single-valued at $p$. 
Then if $L: \reals^d \times \Y \to \reals$ indirectly elicits $\gamma$, then there is a flat $F$ containing $p$, with $F \cap \simplex \subseteq \gamma_r$, and $\codim(F) \leq d$.
% COmmented 06.03.20
%	If there is a convex surrogate loss $L : \reals^d \times\Y \to \reals$ indirectly eliciting $\gamma$, then for all $p \in \simplex$ and $r \in \R$ such that $r \in \gamma(p)$, there is a flat $F \subseteq \affhull(\simplex)$ of codimension at most $d$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$.
\end{theorem}
\begin{proof}
	Let $\Gamma := \prop{L}$ (nonempty by assumption) and suppose $(L, \psi)$ indirectly elicits $\gamma$, so that for all $u \in \reals^d$, we have $\Gamma_u \subseteq \gamma_{\psi(u)}$.
	Fix $p \in \simplex$ and $r \in \gamma(p)$.
  \jessie{Show that there is some $u \in \Gamma(p)$ such that $\psi(u) = r$.}
	We will show $F \cap \simplex \subseteq \Gamma_u$; the result then follows from $\Gamma_u \subseteq \gamma_r$.
	As $L$ is convex and elicits $\Gamma$, we have $u \in \Gamma(p) \iff \vec 0 \in \partial \exploss{L}{u}{p} \iff \vec 0 \in \oplus_y p_y \partial L(u,y)$, where $\oplus$ denotes the Minkowski sum~\citep[Theorem 4.1.1]{hiriart2012fundamentals}.
  % (Observe that in order for $L$ to be convex, $L(\cdot, y)$ must be convex for all $y \in \Y$.)
